---
title: "Sensitivity and Specificity"
author: "Derek Sonderegger"
date: "`r Sys.Date()`"
#output: rmarkdown::word_document
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Sensitivity and Specificity}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE
)
```

```{r, warning=FALSE, message=FALSE}
# library(devtools)
# install_github('dereksonderegger/BurkPx')  # only need to do this once...
library(BurkPx)
library(ggplot2)
```


# Human ROC analysis on Human models
*Note to self:  I need to ask Erik for a description of the human trials we used so as to 
document what data we used to fit the models.*

The models were fit using all of the time points. If an individual had multiple time points,
we only used one time point (randomly selected). The models we are using were fit using 
cross-validation so they shouldn't be overfitting
the data too much, but in the analysis we are about to do, we are using the _same data_ as
we used to create the models.  Ideally, we would have a separate validation set at some point.

## Model Selection
To assess how well the models work, we will look at the Area Under the Curve (AUC) for the 
Reciever-Operator Curve (ROC). To generate this by week, We take the human data and split it 
into Healthy, Week 1, Week 2, etc. For each of the Weeks, we calculate probability of infection 
for the the Healthy and Infected groups using the six different human models. 

```{r}
#################################################
##            ROC for each model               ##
#################################################
AUC_results <- NULL
for(time in c('Week1','Week2','Week3', 'Week4+')){
  for(method in c('LASSO', 'Ridge')){
    for(type in c('IgM', 'IgG', 'IgGM')){
      if( type %in% c('IgM', 'IgG') ){
        df <- Human_BurkPx %>%
          filter(Type == type) %>% filter(TimeGroup %in% c('Healthy', time)) %>%
          spread(Antigen, Value) %>%
          group_by(PatientID) %>% sample_n(1) %>%  # Use only 1 replicant from each patient
          group_by() %>% select(-PatientID, -TimeGroup, -Type, -Rep) %>% complete()
      }else{
        df <- Human_BurkPx %>%
          group_by(PatientID, Type, Rep) %>%
          unite( 'Antigen', Type, Antigen ) %>%
          spread(Antigen, Value) %>%
          group_by(PatientID) %>% sample_n(1) %>%  # Use only 1 replicant from each patient
          group_by() %>% select(-PatientID, -TimeGroup, -Rep)
      }

      df$p <- predict(models[[str_c('Human_',type,'_',method)]], 
                      newdata=df, na.action = na.pass) %>% 
              as.vector()
      temp <- pROC::roc(Status ~ p, data=df)
      AUC_results <- AUC_results %>%
        rbind( data.frame(Type = type, TimeGroup=time, method=method, AUC=pROC::auc(temp) ) )
    }
  }
}
#AUC_results %>% arrange(TimeGroup, desc(AUC))
```



```{r, fig.width=7}
ggplot(AUC_results, aes(x=TimeGroup, y=AUC, color=method)) +
  geom_point() + facet_grid(.~Type) + geom_line(aes(x=as.numeric(TimeGroup)))
```

From this analysis, it is clear that for the first week, we need both the IgG and IgM sereologies. It also is
clear that the LASSO is working better than Ridge Regression and that IgG is working better than IgM. However the
best performance is by the IgGM data which uses both IgG and IgM observation values.


## How many covariates are used?
```{r}
# the models object is a list of all the models I created.  The naming convention
# follows the following convention: Human_IgGM_LASSO.  Species_AntigenList_Method

# Which covariates are used in the IgG LASSO model?
temp <- coef( models$Human_IgG_LASSO ) 
data.frame(Antigen=rownames(temp), Coef=as.vector(temp)) %>%
  filter(Antigen != '(Intercept)', Coef != 0 ) 

# Which covariates are used in the IgGM LASSO model?
temp <- coef( models$Human_IgGM_LASSO ) 
data.frame(Antigen=rownames(temp), Coef=as.vector(temp)) %>%
  filter(Antigen != '(Intercept)', Coef != 0 ) 

# Of those, which are used in both IgG and IgM?
data.frame(Antigen=rownames(temp), Coef=as.vector(temp)) %>%
  filter(Antigen != '(Intercept)', Coef != 0 ) %>%
  separate(Antigen, into=c('Type','Antigen'), extra='merge') %>%
  group_by(Antigen) %>% arrange(Antigen) %>% count()



```


